\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage[a4paper, margin=1.0in]{geometry}
\usepackage{graphicx}

\begin{document}

\begin{center}
\textbf{ELEC-H-473 : Microprocessor Architectures\\ Lab 5 : dsPIC33 1/2\\Arnaud Rosette, Simon Picard}
\end{center}

\subsubsection*{Question 1}
It is a Modified Harvard architecture with 16-bit wide data path and 24-bit wide instructions.\\
There are 83 base instructions (mostly 1 word/1 cycle).\\
The number of register is sixteen and these registers are 16-bit General Purpose Registers. There are also two 40-bit accumulators. The addressing modes are immediate, direct and indirect.

\subsubsection*{Question 2}
A microprocessor contains only a CPU while the microcontroller contains a CPU, some RAM some ROM and other peripherals all embedded in a single chip.\\
A microcontroller is a standalone component which is not the case for a microprocessor.

\subsubsection*{Question 3}
The addressing modes for data are : file register, register direct, register indirect and immediate.

\subsubsection*{Question 4}
\begin{itemize}
\item \textbf{File register addressing} : MOV 0x27FE, W0 (move data stored at 0x27FE to W0)
\item \textbf{Register direct addressing} : EXCH W2,W3 (exchange the content of registers W2 and W3)
\item \textbf{Register indirect addressing} : MOV [W0],[W13] (move data stored at address in W0 to address in W13)
\item \textbf{Immediate addressing} : ADD.B $\#$0x10,W0 (add 0x10 to W0)
\end{itemize}

\subsubsection*{Question 5}
The size of an instruction is 24 bits.\\
Yes, IR is connected to a 24 bits bus.

\subsubsection*{Question 6}
There is 83 instructions so the opcode should be on 7 bits. There is 16 registers so their address are held on 4 bits. Then each instructions might have a different number of parameters but the maximum is 4, if the instruction use an immediate, it is bound by the remaining bits, e.g an instruction using two registers and an immediate : the maximum value of the immediate is $2^{(24-7-4-4)}$

\subsubsection*{Question 7}
This header file provides a C interface for I/O to programmers. It makes real registers usable from C code.

\subsubsection*{Question 8}
The volatile keyword is used in this header file because some of the variables in this file are real registers. So the value of these variables could change unexpectedly (the value can be changed by an interrupt service routine). The volatile keyword tells the compiler to do not make optimizations.

\subsubsection*{Question 9}
\begin{itemize}
 \item \textbf{char, signed char} : -128 to 127
 \item \textbf{unsigned char} : 0 to 255
 \item \textbf{short, signed short} : -32768 to 32767
 \item \textbf{unsigned short} : 0 to 65535
 \item \textbf{int, signed int} : -32768 to 32767
 \item \textbf{unsigned int} : 0 to 65535
 \item \textbf{long, signed long} : -$2^{31}$ to $2^{31}$-1
 \item \textbf{unsigned long} : 0 to $2^{32}$-1
 \item \textbf{long long**, signed long long**} : -$2^{63}$ to $2^{63}$-1
 \item \textbf{unsigned long long**} : 0 to $2^{64}$-1
 \item \textbf{float} : $-3,40282326*10^{38}$ to $3,40282326*10^{38}$ (normalized)
 \item \textbf{double*} : $-3,40282326*10^{38}$ to $3,40282326*10^{38}$ (normalized)
 \item \textbf{long double} : $-1,7976931*10^{308}$ to $1,7976931*10^{308}$ (normalized)
\end{itemize}

\subsubsection*{Question 10}
The integer types are signed by default.

\subsubsection*{Question 11}
There is no assumption anymore in the fact that integers are signed or not by default. The programmer has to specify if he wants to use a signed or an unsigned integer. Furthermore, the number of bits used to represent a type is specified in the name of the type. It makes the code less dependent of the processor.

\subsubsection*{Question 12}
No answer needed.

\subsubsection*{Question 13}
The variables are stored in the stack.

\subsubsection*{Question 14}
The variables are all initialized by moving an immediate value in one or several registers and then the value of  this (or these) register(s) is stored in the memory. The types contained in 8 bits use two mov.b instructions in order to initialize a variable, those which are contained in 16 bits use two mov.w instructions, those which are contained in 32 bits use four mov.w instructions and those which are contained in 64 bits use eight mov.w instructions. The indirect addressing mode is used here, a register (here w14) contains a base address and then an offset is used in order to store variables at different locations in the memory.

\subsubsection*{Question 15}
\begin{verbatim}
INT8U a = 180; // 45 cycle
INT8S b = -105; // 47 cycle
INT16U c = 0x3210; // 49 cycle
INT16S d = -30000; // 51 cycle
INT32U e = 129856; // 55 cycle
INT32S f = -128423; // 59 cycle
INT64U g = 0xFEDCBA9876543210; // 67 cycle
INT64S h = -0x123456789ABCDF0; // 75 cycle
FP32 i = 1.27E+33; // 79 cycle
FP64 j = 8.76543E-18; // 87 cycle
glob1=132; // 88 cycle
\end{verbatim}

\subsubsection*{Question 16}
All the variable but glob1 are load at the first step of the debugger, variable g and h have the same hexadecimal value while not the same decimal value.\\
glob1 is set once the debugger execute its line.

\subsubsection*{Question 17}
The values are temporarily stored in registers from WREG4 to WREG7, for INT16U it uses one register, 2 for INT32U and 4 for INT64U. The higher the number of the register the higher the importance in the number, i.e, WREG7 is the MSB in a INT64U, so to read a INT64U in register you have to concat WREG7, WREG6, WREG5, WREG4.
In the register you can't detect signed number directly.

\subsubsection*{Question 18}
The warnings are about unused variable, irrelevant for this test. There is also a warning about the assignation of glob1 because, it is a cast from an integer to an address, it is not careful to do so.

\subsubsection*{Question 19}
This global variable is defined at line 4, out of the main function, it is set to the address 0x10 which is that of the w8 register.\\
The value is initialized at line 23, at the end of the main function.

\subsubsection*{Question 20}
By using the register keyword, we can declare a variable which will be a register. So when we manipulate this variable, we manipulate a real register. The only difference with directly manipulating the real register is that the content of the register can be interpreted. For example, the content of the register is interpreted differently if the type of the variable is signed or unsigned. The syntax of this keyword is : register <type of the variable> <name of the variable> asm (<name of the register>);

\subsubsection*{Question 21}
First problem : too big integer for unsigned integer, therefore truncated.\\
Too big integer for signed integer.\\
Too big double assigned to infinity.\\
Too small double assigned to 0.\\
The negative value are stored using the ones' complement.

\subsubsection*{Question 22}
The assignation works with cast, the cast from a smaller size to a bigger is no problem, in the other way it takes the LSB. A cast from an int to a float works well but might loose some precision.

\subsubsection*{Question 23}
Firstly, the value of the variable are load in the register, if the size of the variable is smaller than the size of the value then the variable will be assigned with the LSB of the value.\\
To assign float value, it uses some temp register to do some additional computation in order to write the correct representation.

\end{document}
